Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-06-16 manuscript.Rmd
sensor comparison done

@@

@@

Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position. For the sake of brevity, we chose to report the best approach, not every possible variation in modeling. However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values. As in past work [@AiraksinenRasanen2020], performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only). We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion. We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best. Finally, we tried modeling each age group separately, but found no benefit to performance by tailoring to the particular age group.

>>>>

<!-- data/reduce, duplicates of the group_split metrics but performed on reductions of sensor features from group_split_sensor_reduce.R -->

```{r metrics-reduce, cache=TRUE, cache.extra=tools::md5sum("data/reduce/reduce_metrics_right_side.csv")}
all_files = list.files("data/reduce", full.names = T, pattern = ".csv")
overall_files = all_files %>% discard(str_detect(., pattern = "class"))
sensor_comp <- read_csv(overall_files, id = "type")
sensor_comp <- sensor_comp %>% mutate(type = str_remove(type, "data/reduce/reduce_metrics_"),
                              type = str_remove(type, ".csv"),
                              type = factor(type,
                              levels = c("all", "left_side", "right_side", "hips", "ankles",
                                         "left_hip","left_ankle", "right_hip", "right_ankle"),
                              labels = c("All", "Left Hip/Ankle", "Right Hip/Ankle", "Both Hips", "Both Ankles", 
                                         "Left Hip", "Left Ankle", "Right Hip", "Right Ankle")),
                              Model = factor(model, levels = c("group", "split", labels = c("Group", "Individual"))))

metrics_sensor <- sensor_comp %>% group_by(Model, type) %>% 
  get_summary_stats(`Overall Accuracy`) %>% 
  select(-(variable:max), -(q1:mad), -se, -ci) %>% 
  pivot_wider(names_from = Model, values_from = c("median", "mean", "sd")) %>% 
  relocate(mean_group, .after = median_group) %>% 
  relocate(sd_group, .after = mean_group) %>% 
  mutate(across(-type, ~ apa_num(.x, digits = 3)))

apa_table(metrics_sensor, col.names = c("Sensors", "Median", "Mean", "SD", "Median", "Mean", "SD"),    col_spanners = list("Group" = c(2, 4), "Individual" = c(5, 7)), midrules = c(1,5),
          caption = "Median, mean, and SD of overall accuracy calculated with different sets of sensor features. The top row shows performance using all features calculated from the four sensors. Rows 2-5 show accuracy using pairs of sensors (left hip and ankle, right hip and ankle, left and right hip, left and right ankle), and rows 6-9 show accuracy using each individual sensor.")
metrics_sensor <- metrics_sensor %>% column_to_rownames("type")
```

@@

Similar to past work [@AiraksinenRasanen2020], overall accuracy decreased when fewer sensors were used. Table \@ref(tab:metrics-reduce) compared overall accuracy for all four sensors (top row), pairs of 2 sensors (rows 2-5), and single sensors (rows 6-9). The highest accuracy was observed when using all four sensors (group model *M* = `r metrics_sensor["All", "mean_group"]`; individual model *M* = `r metrics_sensor["All", "mean_split"]`), and the lowest when using only a single sensor (left ankle group model *M* = `r metrics_sensor["Left Ankle", "mean_group"]`; right ankle individual model *M* = `r metrics_sensor["Right Ankle", "mean_split"]`). Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and hip has group model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Left Hip/Ankle", "mean_split"]`, only 2.2% and 1.5% worse than using all four sensors. Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_group"]` and individual model accuracy of *M* = `r metrics_sensor["Both Ankles", "mean_split"]`, 12.7% and 7.2% worse than using all four sensors.

@@

As in past work [@AiraksinenRasanen2020], performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only). Small decrements and accuracy were seen for some pairs (particularly when using the left ankle and hip), but decrements for other pairs (both ankles) or individual sensors were substantial. Using a hip and ankle pair instead of four sensors would be more cost effective without a large decrease in accuracy, but in some applications the increased accuracy might be worth the cost. Beyond group and individual models and sensor pairings, there were many degrees of freedom in our choice of how best to model body position. For the sake of brevity, we chose to report the best approach, not every possible variation in modeling. However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values. We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion. We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best. Finally, we tried modeling each age group separately, but found no benefit to performance by tailoring to the particular age group.

### 2023-06-16 manuscript.tex
sensor comparison done

\hypertarget{sensor-configuration}{%
\subsubsection{Sensor Configuration}\label{sensor-configuration}}

@@

@@

Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position. For the sake of brevity, we chose to report the best approach, not every possible variation in modeling. However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values. As in past work \autocite{AiraksinenRasanen2020}, performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only). We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion. We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best. Finally, we tried modeling each age group separately, but found no benefit to performance by tailoring to the particular age group.

>>>>

@@

Similar to past work \autocite{AiraksinenRasanen2020}, overall accuracy decreased when fewer sensors were used. Table \ref{tab:metrics-reduce} compared overall accuracy for all four sensors (top row), pairs of 2 sensors (rows 2-5), and single sensors (rows 6-9). The highest accuracy was observed when using all four sensors (group model \emph{M} = 0.846; individual model \emph{M} = 0.916), and the lowest when using only a single sensor (left ankle group model \emph{M} = 0.667; right ankle individual model \emph{M} = 0.819). Some sensor pairs showed comparable accuracy to using all four sensors: Left ankle and hip has group model accuracy of \emph{M} = 0.824 and individual model accuracy of \emph{M} = 0.901, only 2.2\% and 1.5\% worse than using all four sensors. Other pairings were less accurate; notably, using both ankles resulted in group model accuracy of \emph{M} = 0.719 and individual model accuracy of \emph{M} = 0.844, 12.7\% and 7.2\% worse than using all four sensors.

@@

As in past work \autocite{AiraksinenRasanen2020}, performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only). Small decrements and accuracy were seen for some pairs (particularly when using the left ankle and hip), but decrements for other pairs (both ankles) or individual sensors were substantial. Using a hip and ankle pair instead of four sensors would be more cost effective without a large decrease in accuracy, but in some applications the increased accuracy might be worth the cost. Beyond group and individual models and sensor pairings, there were many degrees of freedom in our choice of how best to model body position. For the sake of brevity, we chose to report the best approach, not every possible variation in modeling. However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values. We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion. We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best. Finally, we tried modeling each age group separately, but found no benefit to performance by tailoring to the particular age group.
